---
title: "HUK Case Study - Model Building"
author: "Chris-Gabriel Islam"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    self-contained: true
---

# Init

```{r setup}
#| output: false
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
library(caret) # for data partition
library(dplyr) # for data manipulation
library(h2o) # for machine learning
```

# Read data

```{r read data}
train_data <- readRDS(here::here("data", "edit", "train_data.rds"))
test_data <- readRDS(here::here("data", "edit", "test_data.rds"))
```

# Improve XGBoost model

Sticking with RMSE as the evaluation metric, we can further improve the XGBoost model by tuning the hyperparameters. We use the `caret` package to tune the hyperparameters. We use a grid search to find the best hyperparameters. We use 5-fold cross-validation to evaluate the model. We use the `train` function to tune the hyperparameters.

```{r tune xgboost}
# Initialize H2O
h2o.init()

# Load Data
train_h2o <- as.h2o(train_data)
test_h2o <- as.h2o(test_data)

# Define Predictors and Target
predictors <- setdiff(names(train_h2o), "target")
target <- "target"

# Define Hyperparameter Grid
hyper_params <- list(
  ntrees = c(50, 100, 200),
  max_depth = c(4, 6, 8),
  learn_rate = c(0.01, 0.05, 0.1),
  colsample_bytree = c(0.7, 0.8, 1.0),
  subsample = c(0.7, 0.8, 1.0),
  min_rows = c(1, 5, 10)
)

# Define Search Criteria (Random Search)
search_criteria <- list(
  strategy = "RandomDiscrete",
  max_models = 20,
  seed = 42
)

# Run Grid Search
grid <- h2o.grid(
  algorithm = "xgboost",
  grid_id = "xgb_grid",
  x = predictors,
  y = target,
  training_frame = train_h2o,
  validation_frame = test_h2o,
  stopping_metric = "RMSE",
  hyper_params = hyper_params,
  search_criteria = search_criteria
)

# Get Best Model (Lowest RMSE)
grid_results <- h2o.getGrid("xgb_grid", sort_by = "RMSE", decreasing = FALSE)
best_model <- h2o.getModel(grid_results@model_ids[[1]])

# Print Best Model Parameters
print(best_model@parameters)

# Compute RMSE
perf <- h2o.performance(best_model, newdata = test_h2o)
rmse <- h2o.rmse(perf)
print(paste("Best Model RMSE:", rmse))
```

# Explain model

```{r explain model}
# Initialize H2O Explain
h2o_explain <- h2o.explain(best_model, test_h2o)
print(h2o_explain)
# Close h2o
h2o.shutdown(prompt = FALSE)
```
